---
format: 
  revealjs:
    slide-number: true
    theme: [default, custom_theme.scss]
    preview-links: auto
    output-file: "index"
    output-ext: "html"
    code-line-numbers: true
    code-overflow: wrap
    embed-resources: true
---

```{r}
#| label: setup
#| include: false
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  out.width = "100%"
)

include_local_figure <- function(data_source) {
  knitr::include_graphics(
    path = here::here(
      "Presentation/Materials",
      data_source
    ),
    error = TRUE
  )
}

library(here)

here::i_am("presentation.qmd")

```

# {data-background-image="./Presentation/Materials/Lecture series-Spring25.png" background-size="900px" background-repeat="no-repeat"}

# {.title}

:::: {.columns}

::: {.column width="5%"}
:::

::: {.column width="20%"}

```{r}
#| label: fossilpol_logo_1
include_local_figure("Logos/FOSSILPOL_logo_600ppi.png")
```
:::

::::

::: {.r-hstack}
::: {data-id="box1" style="background: #d55288; width: 120px; height: 15px; margin: 0px;"}
:::

::: {data-id="box2" style="background: #415280; width: 550px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: #ff9714; width: 120px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: #eb5f56; width: 120px; height: 15px; margin: 0px;"}
:::
:::

::: {.text-bold}
The Workflow To Process [Global]{.text-background-purple} Palaeoecological Data Of Fossil Pollen For [Vegetation]{.text-background-pink}-Based Macroecological [Synthesis]{.text-background-orange}
:::

<br>

::: {.text-right}
Ondřej Mottl[🔗](https://ondrejmottl.github.io/)

27.3.2025
:::


## {.bg-black}

:::: {.columns}

::: {.column width="50%"}

<br>
<br>

```{r}
#| label: earth image
include_local_figure("AI_generated/Earth.jpeg")
```

:::

::: {.column width="50%" .text-center .center-vertical}

7/8 [global]{.text-background-purple} Earth system [boundaries]{.text-background-pink} are already [exceeded]{.text-background-orange} 

::: {.text-smaller .text-color-grey}
(Rockström et al. (2023))
:::

:::

::::

::: {.footer}
Image generated by Bing
:::

## {background-image="./Presentation/Materials/Pexels/pexels-justus-menke-3490295-5214145.jpg" background-opacity="0.65"}

<br>
<br>
<br>
<br>

:::: {.blockquote .text-center .bg-white-50 .text-bold .text-larger} 
Humans altered [> 75%]{.text-background-purple} of ice-free land 
::::

::: {.footer}
Ellis, E. C. & Ramankutty, N. (2008)

Image by Justus Menke from Pexels
:::

## When?

<br>
<br>

:::: {.columns}

::: {.column width="33%" .fragment}
```{r}
#| label: when-1
include_local_figure("AI_generated/prehistoric_man.jpeg")
```
:::

::: {.column width="33%" .fragment}

```{r}
#| label: when-2
include_local_figure("AI_generated/ricefield.jpeg")
```
:::

::: {.column width="33%" .fragment}

```{r}
#| label: when-3
include_local_figure("AI_generated/industrialization.jpeg")
```

:::

::::

::: {.footer}
Images created by Bing 
:::

## 🙋[Hi!]{.text-background-purple} I am [Ondřej Mottl]{.text-background-orange}

<br>

::: {.text-center}
Assistant Professor at [Charles University](https://cuni.cz/UK-1.html)

Head of 🧑‍💻[Laboratory of Quantitative Ecology](https://ondrejmottl.github.io/lab/about_the_lab.html)

:::

<br>

:::: {.columns}

::: {.column width="60%"}
<br>

* 📧 Ondrej.mottl(at)gmail.com
* 🦋 [ondrejmottl.bsky.social](https://bsky.app/profile/ondrejmottl.bsky.social)
* 🐱 [Github OndrejMottl](https://github.com/OndrejMottl)
* 🆔 [ORCID 0000-0002-9796-5081](https://orcid.org/0000-0002-9796-5081)
* 🌐 [bit.ly/ondrej_mottl](https://bit.ly/ondrej_mottl)
:::

::: {.column width="35%"}
```{r}
#| label: head photo
include_local_figure("About/head_photo.png")
```
:::

::::

## Palaeoecology

<br>

:::: {.columns}

::: {.column width = "50%"}

<br>
<br>
<br>
<br>
<br>

"[Fossil pollen]{.text-background-purple } as a window to the past [vegetation]{.text-background-orange} patterns"
:::

::: {.column width = "50%"}

```{r}
#| label: palaeoecology
include_local_figure("Ai_generated/fossilpollen.jpeg")
```
:::

::::

::: {.footer}
Image created by Bing 
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme 1
include_local_figure("Visualisation-fossilpollen_extraction/scheme_1.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme 2
include_local_figure("Visualisation-fossilpollen_extraction/scheme_2.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme 3
include_local_figure("Visualisation-fossilpollen_extraction/scheme_3.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme full
include_local_figure("Visualisation-fossilpollen_extraction/scheme_full.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::


## {background-image="./Presentation/Materials/R_generated/Neotoma_spatial_data_full-2025-03-13.png" background-size="1500px" background-repeat="no-repeat" background-opacity="0.65"}

<br>
<br>
<br>

:::: {.columns}

::: {.column width = "20%"}
:::

::: {.column width = "60%"}

::: {.blockquote .text-center .bg-white-50}

::: {.text-bold .r-fit-text .text-color-purple}
\>6500
:::

fossil pollen timeseries (records)

:::

:::

::::

<br>

:::: {.blockquote .text-center .fragment  .bg-white} 
To uncover [precise insights]{.text-background-purple} from paleo-data, we need a [bias-reducing tool]{.text-background-pink} for seamless [data compilation]{.text-background-orange}
::::


::: {.footer}
Fossil pollen datasets accessed from Neotoma on 2025-03-13
:::

# [This presentation]{.bold .r-fit-text} {.text-center}

<br>

:::: columns
::: {.column width="50%"}
[😸 Code on GitHub](https://github.com/HOPE-UIB-BIO/FOSSILPOL-example-Scandinavia/)
:::

::: {.column width="50%"}
[🖥️ Slides](https://hope-uib-bio.github.io/FOSSILPOL-example-Scandinavia/)
:::
::::

[![](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)

Author: [Ondrej Mottl](https://ondrejmottl.github.io/)