---
format: 
  revealjs:
    slide-number: true
    theme: [default, custom_theme.scss]
    preview-links: auto
    output-file: "index"
    output-dir: "docs"
    output-ext: "html"
    logo: "../Presentation/Materials/Logos/FOSSILPOL_logo_600ppi.png"
    code-line-numbers: true
    code-overflow: wrap
    embed-resources: true
    center-title-slide: false
---

```{r}
#| label: setup
#| include: false
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.align = "center",
  out.width = "100%",
  dpi = 300,
  fig.align = "center"
)

include_local_figure <- function(data_source) {
  knitr::include_graphics(
    path = here::here(
      "Presentation/Materials",
      data_source
    ),
    error = TRUE
  )
}

library(here)

here::i_am("Presentation/presentation.qmd")

```

# 

```{r}
#| label: lecture_intro
include_local_figure("Lecture series-Spring25.png")
```

# 

:::: {.columns}

::: {.column width="5%"}
:::

::: {.column width="20%"}

```{r}
#| label: fossilpol_logo_1
include_local_figure("Logos/FOSSILPOL_logo_600ppi.png")
```
:::

::::

::: {.r-hstack}
::: {data-id="box1" style="background: #d55288; width: 120px; height: 15px; margin: 0px;"}
:::

::: {data-id="box2" style="background: #415280; width: 550px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: #ff9714; width: 120px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: #eb5f56; width: 120px; height: 15px; margin: 0px;"}
:::
:::

::: {.text-bold}
The Workflow To Process [Global]{.text-background-purple} Palaeoecological Data Of Fossil Pollen For [Vegetation]{.text-background-pink}-Based Macroecological [Synthesis]{.text-background-orange}
:::

<br>

::: {.text-right}
Ondřej Mottl[👤](https://ondrejmottl.github.io/)

27.3.2025

[Palaeoverse Lecture Series](https://palaeoverse.org/lectures)
:::


## {.bg-black}

:::: {.columns}

::: {.column width="50%"}

<br>
<br>

```{r}
#| label: earth image
include_local_figure("AI_generated/Earth.jpeg")
```

:::

::: {.column width="50%" .text-center .center-vertical}

7/8 [global]{.text-background-purple} Earth system [boundaries]{.text-background-pink} are already [exceeded]{.text-background-orange} 

::: {.text-smaller .text-color-grey}
(Rockström et al. (2023))
:::

:::

::::

::: {.footer}
Image generated by Bing
:::

## {background-image="../Presentation/Materials/Pexels/pexels-justus-menke-3490295-5214145.jpg" background-opacity="0.65"}

<br>
<br>
<br>
<br>

:::: {.blockquote .text-center .bg-white-50 .text-bold .text-larger} 
Humans altered [> 75%]{.text-background-purple} of ice-free land 
::::

::: {.footer}
Ellis, E. C. & Ramankutty, N. (2008)

Image by Justus Menke from Pexels
:::

## When?

<br>
<br>

:::: {.columns}

::: {.column width="33%" .fragment}
```{r}
#| label: when-1
include_local_figure("AI_generated/prehistoric_man.jpeg")
```
:::

::: {.column width="33%" .fragment}

```{r}
#| label: when-2
include_local_figure("AI_generated/ricefield.jpeg")
```
:::

::: {.column width="33%" .fragment}

```{r}
#| label: when-3
include_local_figure("AI_generated/industrialization.jpeg")
```

:::

::::

::: {.footer}
Images created by Bing 
:::

## 🙋[Hi!]{.text-background-purple} I am [Ondřej Mottl]{.text-background-orange}

<br>

::: {.text-center}
Assistant Professor at [Charles University](https://cuni.cz/UK-1.html)

Head of 🧑‍💻[Laboratory of Quantitative Ecology](https://ondrejmottl.github.io/lab/about_the_lab.html)

:::

<br>

:::: {.columns}

::: {.column width="60%"}
<br>

* 📧 Ondrej.mottl(at)gmail.com
* 🦋 [ondrejmottl.bsky.social](https://bsky.app/profile/ondrejmottl.bsky.social)
* 🐱 [Github OndrejMottl](https://github.com/OndrejMottl)
* 🆔 [ORCID 0000-0002-9796-5081](https://orcid.org/0000-0002-9796-5081)
* 🌐 [bit.ly/ondrej_mottl](https://bit.ly/ondrej_mottl)
:::

::: {.column width="35%"}
```{r}
#| label: head photo
include_local_figure("About/head_photo.png")
```
:::

::::

## Palaeoecology

<br>

:::: {.columns}

::: {.column width = "50%"}

<br>
<br>
<br>
<br>
<br>

"[Fossil pollen]{.text-background-purple } as a window to the past [vegetation]{.text-background-orange} patterns"
:::

::: {.column width = "50%"}

```{r}
#| label: palaeoecology
include_local_figure("Ai_generated/fossilpollen.jpeg")
```
:::

::::

::: {.footer}
Image created by Bing 
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme 1
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-fossilpollen_extraction/scheme_1.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme 2
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-fossilpollen_extraction/scheme_2.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme 3
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-fossilpollen_extraction/scheme_3.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::

## Fossil pollen dataset [(record)]{.text-color-grey}

<br>

```{r}
#| label: fp_scheme full
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-fossilpollen_extraction/scheme_full.png")
```

::: {.footer}
Artwork by Milan Teunissen van Manen (@MilanTvM).
:::


## {background-image="../Presentation/Materials/R_generated/Neotoma_spatial_data_full-2025-03-13.png" background-size="1500px" background-repeat="no-repeat" background-opacity="0.65"}

<br>
<br>
<br>

:::: {.columns}

::: {.column width = "20%"}
:::

::: {.column width = "60%"}

::: {.blockquote .text-center .bg-white-50}

::: {.text-bold .r-fit-text .text-color-purple}
\>6500
:::

fossil pollen timeseries (records)

:::

:::

::::

<br>

:::: {.blockquote .text-center .fragment  .bg-white} 
To uncover [precise insights]{.text-background-purple} from paleo-data, we need a [bias-reducing tool]{.text-background-pink} for seamless [data compilation]{.text-background-orange}
::::


::: {.footer}
Fossil pollen datasets accessed from Neotoma on 2025-03-13
:::

## Palaeoecological dataset compilation

<br>

:::: {.columns}

::: {.column width = "70%"}

```{r}
#| label: compilation_1
include_local_figure("Visualisation-compilation/compilation_1.png")
```
:::

::: {.column width = "30%"}
:::

::::

## Palaeoecological dataset compilation

<br>

:::: {.columns}

::: {.column width = "70%"}

```{r}
#| label: compilation_2
include_local_figure("Visualisation-compilation/compilation_2.png")
```
:::

::: {.column width = "30%"}
:::

::::

## Palaeoecological dataset compilation

<br>

:::: {.columns}

::: {.column width = "70%"}

```{r}
#| label: compilation_3
include_local_figure("Visualisation-compilation/compilation_3.png")
```
:::

::: {.column width = "30%"}
:::

::::

## Palaeoecological dataset compilation

<br>

:::: {.columns}

::: {.column width = "70%"}

```{r}
#| label: compilation_4
include_local_figure("Visualisation-compilation/compilation_4.png")
```
:::

::: {.column width = "30%"}
:::

::::

## Palaeoecological dataset compilation

<br>

:::: {.columns}

::: {.column width = "70%"}

```{r}
#| label: compilation_5
include_local_figure("Visualisation-compilation/compilation_5.png")
```
:::

::: {.column width = "30%"}
:::

::::


## Palaeoecological dataset compilation

<br>

:::: {.columns}

::: {.column width = "70%"}

```{r}
#| label: compilation_full
include_local_figure("Visualisation-compilation/compilation_full.png")
```
:::

::: {.column width = "30%" .incremental}

- Macro-scale data analysis
- Project specific
- Data quality control and reproducibility

:::

::::

# 

:::: {.columns}

::: {.column width = "50%"}
```{r}
#| label: fossilpol_logo_2
include_local_figure("Logos/FOSSILPOL_logo_600ppi.png")
```
:::

::: {.column width = "50%" .center-vertical .text-larger .text-center}
The workflow to process global [palaeoecological data]{.text-background-purple} of fossil pollen for [vegetation]{.text-background-pink }-based [macroecological]{.text-background-coral} [synthesis]{.text-background-orange}
:::

::::

## Example project {background-image="../Presentation/Materials/Pexels/pexels-damir-31222883.jpg" background-opacity="0.65"}

<br>

:::: {.blockquote .text-center .bg-white-50 .text-bold .text-larger}
What were the patterns of past [vegetation]{.text-background-purple} change in [Scandinavia]{.text-background-pink} during [the last 8,000 years]{.text-background-orange}?
::::

::: {.footer}
Image by Damir K. from Pexels
:::

## 

<br>

```{r}
#| label: workflow full
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_full.png")
```

## 

<br>

```{r}
#| label: workflow_clean
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_clean.png")
```

##

<br>

```{r}
#| label: workflow_0
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_0.png")
```

## [Inputs]{.text-color-white} {background-image="../Presentation/Materials/Banners/Banner_1.png" background-size="100% 20%" background-position="top left" background-repeat="no-repeat"}

<br>

:::: {.columns}
::: {.column width="10%"}

<br>

```{r}
#| label: icon_config
include_local_figure("Icons/icon_config_file.png")
```
:::

::: {.column width="45%" .incremental}

<br>

### Config file

- User-defined criteria guided by the research questions
- Control the behaviour of the process
- Reproducible


:::

::: {.column width="45%"}
```{r}
#| label: puppeteer
include_local_figure("AI_generated/puppeteer.jpeg")
```
:::

::::

::: {.footer}
Image created by Bing
:::

##

<br>

```{r}
#| label: workflow_0b
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_0.png")
```


## 

<br>

```{r}
#| label: workflow_1
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_1.png")
```

## [SOURCING]{.text-color-white} {background-image="../Presentation/Materials/Banners/Banner_2.png" background-size="100% 20%" background-position="top left" background-repeat="no-repeat"}

<br>

:::: {.columns}

::: {.column width="60%" .incremental}

- Main source - [Neotoma]{.text-background-blue}
- Other (private) - possible
- Spatial delimitation
- …

```{r}
#| label: logo_neotoma
include_local_figure("Logos/neotoma_logo.png")
```
:::

::: {.column width="40%" .fragment .center-vertical}

<br>
<br>

```{r}
#| label: map_1
include_local_figure("R_generated/sc_map_sourcing.png")
```
:::

::::


## 

<br>

```{r}
#| label: workflow_2
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_2.png")
```

## 

<br>

```{r}
#| label: workflow_3
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_3.png")
```

## [CHRONOLOGIES]{.text-color-white} {background-image="../Presentation/Materials/Banners/Banner_3.png" background-size="100% 20%" background-position="top left" background-repeat="no-repeat"}

<br>

:::: {.columns}

::: {.column width="10%"}

```{r}
#| label: icon_bchron
include_local_figure("Logos/Bchron_logo.png")
```

:::

::: {.column width="50%" .incremental}

- Re-estimate temporal position 
  - [Bayesian]{.text-background-purple} probabilistic methododlogy
- Calibration curves
  - by geography
  - postbomb
- [Chronology control tables]{.text-background-purple}
  - Chron.c.point types
  - Number of dates
- …

:::

::: {.column width="40%" .fragment .center-vertical}

<br>
<br>

```{r}
#| label: map_2
include_local_figure("R_generated/sc_map_chronologies.png")
```
:::

::::

## 

```{r}
#| label: workflow_4
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_4.png")
```

##

```{r}
#| label: workflow_5
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_5.png")
```

## [HARMONISATION]{.text-color-white} {background-image="../Presentation/Materials/Banners/Banner_4.png" background-size="100% 20%" background-position="top left" background-repeat="no-repeat"}

<br>

:::: {.columns}

::: {.column width="75%"}
- Reduce the effect of [taxonomic uncertainty]{.text-background-pink} and nomenclatural complexity
- Using [harmonisation table]{.text-background-pink}(s)

<br>

:::: {.columns .fragment fragment-index=4}

::: {.column width="15%"}
:::

::: {.column width="70%"}
```{r}
#| label: harm_paper_preview
include_local_figure("Printscreens/harm_paper_preview.png")
```

[![](https://img.shields.io/badge/DOI-10.1016%2Fj.revpablo.2023.104989-yellow)](https://doi.org/10.1016/j.revpalbo.2023.104989)
:::

::: {.column width="15%"}
:::

::::

:::


::: {.column width="5%"}
:::

::: {.column width="20%"}

::: {.fragment fragment-index=2}

```{r}
#| label: flowers_diff
include_local_figure("Icons/flowers_diff.png")
```
:::

```{r}
#| label: icon_harm_table
include_local_figure("Icons/icon_harm_table.png")
```

::: {.fragment fragment-index=3}
```{r}
#| label: flowers_same
include_local_figure("Icons/flowers_same.png")
```
:::

:::

::::

::: {.footer}
Flower icons created by Smashicons - Flaticon
:::

##

```{r}
#| label: workflow_6
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_6.png")
```

##

```{r}
#| label: workflow_7
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_7.png")
```

## [FILTERING]{.text-color-white} {background-image="../Presentation/Materials/Banners/Banner_5.png" background-size="100% 20%" background-position="top left" background-repeat="no-repeat"}

<br>

:::: {.columns}

::: {.column width="60%" .incremental}
- selection of samples and datasets
- [increase data quality]{.text-background-coral}
- Optional
  - Pollen sum
  - Age limits
  - Age extrapolation
  - Min N samples
  - …
:::

::: {.column width="40%" .fragment .center-vertical}

<br>
<br>

```{r}
#| label: map_3
include_local_figure("R_generated/sc_map_filtering.png")
```
:::
::::

## Scandinavian example

<br>
<br>

```{r}
#| label: map_overview
include_local_figure("R_generated/sc_map_all.png")
```

::: {.fragment}

```{r}
#| label: data_quality
include_local_figure("R_generated/sc_map_increase_quality.png")
```
:::

## 

```{r}
#| label: workflow_8
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_8.png")
```

##

```{r}
#| label: workflow_fullb
#| out.height: 100%
#| out.width: NULL
include_local_figure("Visualisation-workflow/workflow_full.png")
```

## [Outputs]{.text-color-white} {background-image="../Presentation/Materials/Banners/Banner_6.png" background-size="100% 20%" background-position="top left" background-repeat="no-repeat"}

<br>

:::: {.columns}

::: {.column width="45%"}

<!-- first row-->
:::: {.columns .fragment fragment-index=1}

::: {.column width="10%"}

```{r}
#| label: icon_dataset_compilation
include_local_figure("Icons/icon_data_compilation.png")
```
:::

::: {.column width="70%"}
**Dataset compilation**

[ - Ready for analyses]{.text-smaller}
:::

::: {.column width="10%"}

```{r}
#| label: logo_R
include_local_figure("Logos/R_logo.png")
```
:::

::::

<!-- second row-->
:::: {.columns .fragment fragment-index=2}

::: {.column width="10%"}

```{r}
#| label: icon_graphical_visualisation
include_local_figure("Icons/icon_graph_visualiasation.png")
```
:::

::: {.column width="70%"}
**Graphical visualisation**

[- Spatio-temporal overview]{.text-smaller}
:::

::: {.column width="10%"}
:::

::::

<!-- third row-->
:::: {.columns .fragment fragment-index=4}

::: {.column width="10%"}

```{r}
#| label: icon_citation
include_local_figure("Icons/icon_citation.png")
```
:::

::: {.column width="70%"}
**Citation metadata**

[- Owners and references of data]{.text-smaller}

[- "Reproducibility bundle"]{.text-smaller}
:::

::: {.column width="10%"}

```{r}
#| label: icon_open_science
include_local_figure("Icons/icon_open_science.png")
```
:::

::::

:::

::: {.column width="55%" .center-vertical .fragment fragment-index=3}

<br>

```{r}
#| label: map_summary
include_local_figure("R_generated/sc_map_summary.png")
```
:::

::::

# FOSSILPOL

:::: {.columns}

::: {.column width="75%" .fragment}

<br>

- Handles all the steps and flags places where the [user]{.text-background-purple} needs to specify their [choices]{.text-background-pink}
:::

::: {.column width="25%"}
```{r}
#| label: qr_fossilpol
include_local_figure("QR_codes/qr_fossilpol.png")
```
:::

::::

:::: {.columns .fragment}

::: {.column width="60%"}
- '*A guide to the processing and standardisation of global palaeoecological data for large-scale syntheses using fossil pollen*' (GEB 2023)

[![](https://img.shields.io/badge/DOI-10.1111%2Fgeb.13693-yellow)](https://doi.org/10.1111/geb.13693)
:::

::: {.column width="5%"}
:::

::: {.column width="35%"}

```{r}
#| label: gep_certificate
include_local_figure("Printscreens/GEB_certifiate.png")
```


:::

::::

## Website

<br>

:::: {.columns}

::: {.column width="40%"}

```{r}
#| label: website_preview
include_local_figure("Printscreens/website_preview.png")
```
:::

::: {.column width="60%" .fragment}

```{r}
#| label: website_doc_preview
include_local_figure("Printscreens/website_doc_preview.png")
```

[websites]{.text-background-orange} - [bit.ly/FOSSILPOL](https://bit.ly/FOSSILPOL)

:::

::::

## This presentation {.text-center .bg-black}

<br>
<br>

[😸 Code on GitHub](https://github.com/HOPE-UIB-BIO/FOSSILPOL-example-Scandinavia/)

[![](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)

<br>

:::: {.columns}

::: {.column width="33%"}
:::

::: {.column width="30%"}
```{r}
#| label: qr_slides
include_local_figure("QR_codes/qr_slides.png")
```
:::

::: {.column width="33%"}
:::

::::

## 

:::: {.columns}

::: {.column width="70%"}

<br>
<br>
<br>

::: {.text-bold}
The Workflow To Process [Global]{.text-background-purple} Palaeoecological Data Of Fossil Pollen For [Vegetation]{.text-background-pink}-Based Macroecological [Synthesis]{.text-background-orange}
:::

:::: {.r-hstack}
::: {data-id="box1" style="background: #d55288; width: 120px; height: 15px; margin: 0px;"}
:::

::: {data-id="box2" style="background: #415280; width: 550px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: #ff9714; width: 120px; height: 15px; margin: 0px;"}
:::

::: {data-id="box3" style="background: #eb5f56; width: 120px; height: 15px; margin: 0px;"}
:::
::::

[27.3.2025]{.text-color-grey}

:::

::: {.column width="30%"}

```{r}
#| label: head_photo2
include_local_figure("About/head_photo.png")
```
:::

::::

<br>

:::: {.columns}

::: {.column width="15%"}

```{r}
#| label: personal_sticker
include_local_figure("About/personal_sticker.png")
```
:::

::: {.column width="55%"}
* 📧 Ondrej.mottl(at)gmail.com
* 🦋 [ondrejmottl.bsky.social](https://bsky.app/profile/ondrejmottl.bsky.social)
* 🌐 [bit.ly/ondrej_mottl](https://bit.ly/ondrej_mottl)
:::

::: {.column width="15%"}
```{r}
#| label: qr_fossilpol2
include_local_figure("QR_codes/qr_fossilpol.png")
```
:::

::: {.column width="15%"}
```{r}
#| label: fossilpol_logo_3
include_local_figure("Logos/FOSSILPOL_logo_600ppi.png")
```
:::

::::